---
title: "NoMovementCleaning"
author: "Brandon Mayer"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(maps)
library(mapdata)
library(ggplot2)
library(geosphere)
library(tidyverse)
library(slider)
```

Load in data and Data Pre-wrangle
  -   Assume the use of pycharm for running the Vence-API. This link can be changed to the "Data" folder in the repository if API is run with reticulate
```{r}
wd <- paste0(dirname(path.expand('~')),"/PycharmProjects/Vence-API/data");(files<-list.files(wd,full.names = T))
sampdata <- read.csv(files[1]) 
sampdata <-sampdata[
  order( sampdata[,4], sampdata[,3] ),
] %>% select(-uuid);sampdata$date<- lubridate::ymd_hms(sampdata$date)
```

Load in collar meta data with associated herd, eartag, or additional information
```{r}
# Select a collar export associated with the appropriate herd. Match the collar export with the specific time period you're intersted in. Exports can be retrieved through the Herd Manager app. Exports should be collected prior to time period of interest. 
file;(Metafiles <-list.files(paste0(dirname(path.expand('~')),"/Box/1.Ruyle_lab/1.Project_Data/Virtual Fencing/CollarManagment/RAWCollarHerd_Exports/"),full.names = T))

Meta<- read_csv(Metafiles[1]) %>% select("collar" = DeviceEUI,EarTag,HerdName) %>% mutate(HerdName = "Small Herd") # hard coded for the large herd file of interest


NMC_uncleaned <- sampdata %>% left_join(Meta, by = "collar") %>% drop_na(HerdName)
NMC_uncleaned <- NMC_uncleaned%>% filter(grepl('GpsLocationExtIndication', messagetype))
NMC_uncleaned <- NMC_uncleaned %>% group_by(collar)%>% arrange(collar,date) %>% ungroup()
# Should have no "NAs" in the EarTag column assuming all collars of interest were joined correctly
NMC_uncleaned[is.na(NMC_uncleaned$EarTag),]
```

Step 1:Removing Distance Errors
  -   Add Distance for each collar group, rbind, mutate the time between gps points, assess the rate of movement as DistTime (m/min)
```{r}
NMC_listed <- split(NMC_uncleaned,NMC_uncleaned$collar)
for (i in 1:length(NMC_listed)) {
  NMC_listed[[i]]$Dist <- c(NA)
  for (j in 1:(nrow(NMC_listed[[i]])-1)) {
    NMC_listed[[i]]$Dist[j+1]<- distm(c(NMC_listed[[i]][j,5][[1]],
                                      NMC_listed[[i]][j,4][[1]]),
                                    c(NMC_listed[[i]][(j+1),5][[1]],
                                      NMC_listed[[i]][(j+1),4][[1]]),
                                    fun = distHaversine)
    }
}
NMC_uncleaned <- data.table::rbindlist(NMC_listed)

# Add a rate of movement
NMC_uncleaned <- NMC_uncleaned %>% group_by(collar)%>% arrange(collar,date) %>%
  mutate(time_diff = difftime(date,lag(date), units='mins'))

NMC_uncleaned <- NMC_uncleaned %>% group_by(collar) %>% mutate(DistTime=Dist/as.numeric(time_diff))
NMC_uncleaned <-NMC_uncleaned[!NMC_uncleaned$DistTime >=84,] # 84 m/min: movement speed of a cow

NMC_listed <- split(NMC_uncleaned,NMC_uncleaned$collar)
for (i in 1:length(NMC_listed)) {
  NMC_listed[[i]]$Dist <- c(NA)
  for (j in 1:(nrow(NMC_listed[[i]])-1)) {
    NMC_listed[[i]]$Dist[j+1]<- distm(c(NMC_listed[[i]][j,5][[1]],
                                      NMC_listed[[i]][j,4][[1]]),
                                    c(NMC_listed[[i]][(j+1),5][[1]],
                                      NMC_listed[[i]][(j+1),4][[1]]),
                                    fun = distHaversine)
    }
}
NMC_uncleaned <- data.table::rbindlist(NMC_listed)

```

Step 2:  rowSum collars within a known area
```{r}
# Function to sum the number of collars within a desired error range
count_return <- function(long, lat, dist) {
  clonglat <- cbind(long,lat)
count <- rowSums(distm(clonglat[1,],clonglat, 
       fun = distHaversine)<= dist) # Average max quantile distance for 8 collars in varying canopy densities was 23m
return(count)
 }

# Distance in this case is set to 20 meters and 
test <-NMC_uncleaned %>% group_by(collar) %>% mutate(ErrorGPS = slide2_dbl(.x = longitude,.y = latitude, ~count_return(.x,.y,20),.after =48))

# Function to identify 10 consecutive points with total median counts exceeding 19 and mark them in a new column as 1 (Dropped collar), 0 (No dropped)
DC_screen <- function(Median) {
count <- ifelse(sum(Median >= 19)>=10,1,0)
return(count)
 }

# Comparing the removed collars between counts
for (l in 1:length(testlist)) {
  testlist[[l]] <- testlist[[l]]%>% group_by(collar) %>% mutate(Median = slide_dbl(ErrorGPS,~median(.x),.before=5, .after = 5))%>% mutate(DCScreen = slide_dbl(ErrorGPS,~DC_screen(.x), .after = 10))%>% group_by(collar,DCScreen) %>% mutate(Start = row_number()==1) %>% group_by(collar)%>% mutate(Final=case_when(Start & DCScreen==0~0,Start & DCScreen==1~1))%>% group_by(EarTag) %>% fill(Final,.direction = "down") 
}
testdrop<- list()
for (l in 1:length(testlist)) {
  testdrop[[l]] <- data.frame("collars" = unique(testlist[[l]][testlist[[l]]$Final ==1,]$collar))
}
testdrop
x <- testlist[[7]]
x <- x[x$EarTag %in% "E80",]

# Smooth counts to adjust for minor GPS errors
NMC_NMcleaned <- test %>% group_by(collar) %>% mutate(Median = slide_dbl(ErrorGPS,~median(.x),.before=5, .after = 5))

NMC_NMcleaned <- test %>% group_by(collar) %>% mutate(DCScreen = slide_dbl(ErrorGPS,~DC_screen(.x), .after = 10))

# Identify the first occasion of the above requirements 
NMC_NMcleaned<- NMC_NMcleaned %>% group_by(collar,DCScreen) %>% mutate(Start = row_number()==1) 

# Differentiate first cases between when the collar was active and when it dropped
NMC_NMcleaned <-NMC_NMcleaned %>% group_by(collar)%>% mutate(Final=case_when(Start & DCScreen==0~0,
                                                             Start & DCScreen==1~1)) 
# fill down from the start points( 0 active collar, 1 dropped collar)
NMC_NMcleaned <-NMC_NMcleaned %>% group_by(EarTag) %>% fill(Final,.direction = "down")

# List of collars that have dropped
dropped <- data.frame("collars" = unique(NMC_NMcleaned[NMC_NMcleaned$Final ==1,]$collar))
FinalNMC_NMcleaned <- x %>% select(-c(Dist,DistTime,time_diff,ErrorGPS,DCScreen,Start))

# Write CSV for use in analysis
write.csv(x, paste0(dirname(path.expand('~')),"/Box/1.Ruyle_lab/1.Project_Data/Virtual Fencing/Data/RAW/",unique(x$HerdName),"_",basename(file)))
write.csv(dropped, paste0(dirname(path.expand('~')),"/Box/1.Ruyle_lab/1.Project_Data/Virtual Fencing/Data/Dropped/",basename(file)))

```

